Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-14 doc/invt.tex
intq.h: change the definition of (K -> kc) for mode cutoff

While adaptive FDS simulations
can often sample the desired distribution,
frequent updates to the bias potential
disrupt the underlying equilibrium sampling
and introduce artificial errors that need to be
removed by gradually decreasing the updating magnitude\cite{
  belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
  zhou2005, morozov2007, zhou2008,
  laio2005, bussi2006, poulain2006, liang2007,
  crespo2010, *atchade2011, *fort2015}.
%One can show that with a constant magnitude,
%$\alpha(t) = \alpha > 0$,
%the distribution collected from
%an adaptive FDS simulation
%is identical to $p_i$ in the long time or asymptotic limit.
%
%However, as the above updates
%change $u_i(t)$ continuously,
%the equilibrium distribution,
%Eq. \eqref{eq:pi_p_V}, no longer holds
%at all times,
%%
%The source of the deviation is two-fold.
%%
%First, since $u_i(t)$ is updated continuously,
%there is a random noise that is proportional
%to $\sqrt \alpha$.
%%
%Second, there is a systematic error
%that comes from the updating dynamics itself,
%as it breaks the Markovian nature
%that underlies Eq. \eqref{eq:pi_p_v}.
%%

@@

In analogy to Eq. \eqref{eq:ET_average},
%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:xt2_sum},
we can write the error after a period, $T$,
as a sum of two contributions,
%
\begin{equation}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T),
  \label{eq:error_tot}
\end{equation}
%
where the residual error
(from the decay of the initial error)
is given by
%
\begin{equation}
  \Err_R(T)
  =
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) },
  \label{eq:error_res}
\end{equation}
%
and the asymptotic error
is given by
%
\begin{align}
  \Err_A(T)
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot y_k^2\bigl( q(t) \bigr) \, dt
  ,
\label{eq:error_asym}
\end{align}
%
where
%
\begin{equation}
  y_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  ,
  \label{eq:uk_def}
\end{equation}
%
and $\Gamma_k$ is computed from
the autocorrelation function
of the histogram fluctuation,
$\kappa_k(t) \equiv \bigl\langle {\tilde \zeta}_k(0)
\, {\tilde \zeta}_k(t) \bigr\rangle$,
as:
%
\begin{equation}
  \Gamma_k
  = \sum_{t = -\infty}^\infty \kappa_k(t)
  = \sum_{t = -\infty}^\infty
  \left\langle {\tilde \zeta}_k(0)
            \, {\tilde \zeta}_k(t) \right\rangle
  .
  \label{eq:Gamma_sum}
\end{equation}
%
%\begin{equation}
%$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
%  \notag
%\label{eq:Gamma_integral}
%\end{equation}
%
Equation \eqref{eq:Gamma_sum} provides a way
of approximately modeling
a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
In Appendix \ref{sec:Gamma_measure},
we discuss a way of estimating $\Gamma_k$
from an FDS simulation with a constant magnitude.
%

@@

\item \label{step:prerun}
Run a preliminary adaptive FDS simulation
%using the updating scheme
with constant updating magnitude, $a_0$.

@@

We used the Metropolis algorithm for configuration sampling\cite{frenkel}.
%
Each step consists of
one MC trial for displacing one of the two special particles,
and another four trials for displacing the other particles.
%
In each trial, the displacement along each dimension is uniform,
with the size given by $0.1/\varrho^*$ to ensure the acceptance ratio is around 50\%,
where $\varrho^*$ is the reduced number density.

The PMF was calculated from $r = 0$ to half of the box dimension
on a grid of size $\Delta r = 0.01$ in reduced units.
%
Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
and the numbers of bins were $500$ and $250$, respectively.
%
We tested the single-bin, Gaussian, and bandpass updating schemes.
%
Each simulation started with an ``equilibration'' phase
of $10^7$ steps with a fixed updating magnitude, $a_0 = 10^{-4}$,
followed by a production phase
with updating magnitude given by
either the inverse-time or optimal schedule.
%
We repeated each simulation multiple times
and report the average.
%
To make sure the same schedule for multiple runs under the same condition,
the optimal schedule was pre-computed from a set of fixed values of
$\Gamma_k$ and $\epsilon_k$,
although they can also be fairly accurately estimated from the data
collected from the equilibration phase of each individual run.

@@

In Fig. \ref{fig:lj_xerr},
we show the error components against the eigenmode index
for the Gaussian updating scheme of width $\sigma = 0.2$.
%
While the Gaussian updating scheme effectively suppressed
the initial errors for short-range (large $k$) modes,
it also slowed down the reduction of errors of these modes.
%
As a result, the final errors of the long-range (small $k$) modes
were larger than those from the single-bin scheme.
%
For the Gaussian updating scheme
the inverse-time schedule tended to reduce the updating magnitude too rapid
leaving a hump on the error profile for the middle-range modes.
%
The optimal schedule improved over the inverse-time schedule
and produced a flatter profile for the final error.
%
It is also possible to roughly
recover the single-bin result by using the correction formula
based on the histogram, Eq. \eqref{eq:UHcorr}.
%
This correction usually improves the accuracy of the small $k$ modes
but not necessarily for the large $k$ modes.

In Fig. \ref{fig:lj_alpha}. 
we show the optimal schedules in a few cases.
%
Generally, the optimal schedule
for the $\varrho^* = 0.8$ case
was closer to the inverse-time schedule
than that for the $\varrho^* = 0.1$ case.
%
This was probably because in the former case
the first mode dominates the total error,
$\Gamma_1 \approx \Gamma$,
and the schedule was roughly optimized for this mode,
resulting in
$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
as $\lambda_1 \approx 1$
[especially when width of the Gaussian is small,
cf. Eq. \eqref{eq:lambda_Gaussian}].
%
Since we have chosen a long equilibration period
for preliminary run with a constant updating magnitude,
the starting magnitude was usually around $a_0/2 = 5\times10^{-5}$
[cf. Eq. \eqref{eq:half_alpha0}].
%
But if the Gaussian is too wide,
such as the $\sigma = 0.6$ case,
the eigenvalues $\lambda_k$ of some modes
became too small to damp out the initial bias,
resulting in nonnegligible $\epsilon_k$,
and much larger initial updating magnitude, $\alpha(0)$
[cf. Eq. \eqref{eq:opt_alpha0}].
%
In this case, our method would fail
because the resulting bias potential would no longer be small.

\begin{table}[h]\footnotesize
  \caption{\label{tab:error_LJ}
    Error of the LJ system.
    The numbers have been averaged over multiple runs,
    and the numbers in parentheses are predicted values;
    $\Err_\mathrm{init.}$
    denotes the error
    at the end of the equilibration phase with a constant magnitude;
    $\Err_\mathrm{final}$
    denotes the error at the end of the production phase under
    either the inverse-time ($1/t$) or optimal schedule (opt.).
    \note{The reference values can be found on the second line of \texttt{verr.log}.}
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Sched. &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    \hline
    \multicolumn{4}{c}{
      $\varrho^* = 0.1$,
      $500$ bins,
      $\Gamma_1 \approx 111$,
      $\Gamma \approx 1.1\times10^3$.
    } \\
    \hline
    single-bin & $1/t$
    & $5.7(5.7)\times10^{-2}$
    & $9.1(11.4)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $7.7(8.9)\times10^{-3}$
    & $6.8(8.2)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $7.6(8.9)\times10^{-3}$
    & $3.2(3.9)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $6.5(7.8)\times10^{-3}$
    & $4.0(4.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $7.0(7.8)\times10^{-3}$
    & $2.5(2.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $5.5(6.6)\times10^{-3}$
    & $3.2(3.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $5.8(6.6)\times10^{-3}$
    & $2.6(3.0)\times10^{-5}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $8.1(8.9)\times10^{-3}$
    & $1.6(1.8)\times10^{-5}$
    \\
    \hline
    \multicolumn{4}{c}{
      $\varrho^* = 0.8$,
      $250$ bins,
      $\Gamma_1 \approx 2.0\times10^4$,
      $\Gamma \approx 2.7\times10^4$
    } \\
    \hline
    single-bin & $1/t$
    & $1.3(1.4)$
    & $2.7(2.7)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $1.3(1.3)$
    & $2.9(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $1.2(1.3)$
    & $2.8(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $1.3(1.2)$
    & $3.4(3.2)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $1.2(1.2)$
    & $3.1(2.9)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $0.9(0.9)$
    & $9.8(8.0)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $1.0(0.9)$
    & $5.8(5.4)\times10^{-3}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $1.3(1.3)$
    & $3.0(2.7)\times10^{-3}$
    \\
    \hline
  \end{tabular}
\end{table}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_xerr.pdf}
  }
  \caption{
    \label{fig:lj_xerr}
    Error components for the FDS simulation on the LJ system
    using the Gaussian updating scheme with width $\sigma = 0.2$.
    %
    The optimal schedule (opt.) is more effective
    in reducing the errors of middle-range modes
    than the inverse-time ($1/t$) schedule.
    The lines are a guide to the eyes.
  }
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_alpha.pdf}
  }
  \caption{
    \label{fig:lj_alpha}
    Optimal schedule of the Gaussian updating scheme
    for the LJ system.
    %
    The starting updating magnitude was usually
    around $a_0/2 = 5\times10^{-5}$
    as given by Eq. \eqref{eq:half_alpha0}.
    But an overly wide Gaussian window ($\sigma = 0.6$)
    resulted in a much larger updating magnitude.
  }
\end{center}
\end{figure}

@@

To begin, we split the entire simulation into $S$ stages,
each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
%
We denote the number of steps in stage $s$ as $T_s$, and
$\sum_{s=1}^S T_s = T$.
%
%\begin{equation}
%  \sum_{s = 1}^S T_s = T
%  ,
%  \label{eq:sumT}
%\end{equation}
%
We assume that
the magnitude follows geometrically,
%
\begin{equation}
  \alpha_s = \alpha_0 \, r^s
  ,
  \label{eq:alphas}
\end{equation}
%
with $0 < r < 1$.
%
%and that the system is properly equilibrated
%under a constant magnitude initially.
%
We wish to find the best distribution of the simulation time
into the $S$ stages
as well as
the optimal number of stages and $r$.

@@

Here we give a method of measuring
the integrals of the autocorrelation functions, $\Gamma_k$'s,
in an adaptive FDS simulation
with a constant updating magnitude.
%
This method has the advantage of not requiring
explicit computation of 
the autocorrelation functions of the eigenmodes.
%
The basic idea is to use Eq. \eqref{eq:xt2_eql}.
%
However,
without knowing the intrinsic distribution, $p^*_i$,
we cannot compute the shifted bias potential, $v_i$.
%
The problem can be avoided by replacing the shift
by the time average of the bias potential.

We first note that the difference, 
$u_i(t) - v_i(t) = \ln(p^*_i/p_i)$, is a constant of time,
so is its Fourier transform,
$\tilde u_k(t) - \tilde v_k(t)$.
%
It follows that
${\tilde u}_{k}$ and ${\tilde v}_{k}$ would also share the same variance.
%
But since the ensemble average of ${\tilde v}_{k}$
is zero, we have
\begin{equation*}
  \operatorname{var} {\tilde u}_{k}
  =
  \operatorname{var} {\tilde v}_{k}
  =
  \left\langle {\tilde v}_{k}^2 \right\rangle
  .
\end{equation*}
%
Finally, by using Eq. \eqref{eq:xt2_eql},
we get
%
\begin{equation}
  \operatorname{var} {\tilde u}_k
  =
  \frac{1}{2} \,
  a_0 \, \Gamma_k \, \lambda_k,
\label{eq:varXt}
\end{equation}
%
where the variance can be computed from trajectory averages.
%
A practical limitation is that Eq. \eqref{eq:varXt}
would fail if some eigenvalue, $\lambda_k$, is close to zero.
So it is best used for the single-bin updating scheme.

@@

For the Gaussian updating kernel,
if the width of the angular variable, $\sigma_\varphi$,
satisfies $\Delta \varphi \ll \sigma_\varphi \ll \pi$,
we can extend
the upper limit of the integrals
to infinity, and
%
\begin{equation}
  \mu(\varphi)
  =
  \frac{            1            }
       { \sqrt{ 2 \pi \sigma_\varphi^2 } }
  %
  \exp\left(
        -\frac{   \varphi^2   }
              { 2 \, \sigma_\varphi^2 }
      \right)
  .
\notag
\end{equation}
%
%
It follows that all eigenvalues are positive\cite{bussi2006}:
%
\begin{align}
  \lambda_k
  &=
  \exp\left(
        -\frac{ k^2 \, \sigma_\varphi^2 }
              {           2             }
      \right)
  %\notag
  %\\
  %&=
  %\exp\left[
  %      -\frac{ \pi^2 }{ 2 }
  %      \left(
  %        \frac{ k - 1 }
  %             {   n   }
  %      \right)^2
  %      \left(
  %        \frac{  \sigma_\varphi }
  %             { \Delta \varphi }
  %      \right)^2
  %    \right]
  .
%\notag
\label{eq:lambda_Gaussian}
\end{align}

>>>>

While adaptive FDS simulations
can often sample the desired distribution,
frequent updates to the bias potential
disrupt the underlying equilibrium sampling
and introduce artificial errors that need to be
removed by gradually decreasing the updating magnitude\cite{
  belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
  zhou2005, morozov2007, zhou2008,
  laio2005, bussi2006, poulain2006, liang2007,
  crespo2010, *atchade2011, *fort2015}.
%One can show that under constant magnitude,
%$\alpha(t) = \alpha > 0$,
%the distribution collected from
%an adaptive FDS simulation
%is identical to $p_i$ in the long time or asymptotic limit.
%
%However, as the above updates
%change $u_i(t)$ continuously,
%the equilibrium distribution,
%Eq. \eqref{eq:pi_p_V}, no longer holds
%at all times,
%%
%The source of the deviation is two-fold.
%%
%First, since $u_i(t)$ is updated continuously,
%there is a random noise that is proportional
%to $\sqrt \alpha$.
%%
%Second, there is a systematic error
%that comes from the updating dynamics itself,
%as it breaks the Markovian nature
%that underlies Eq. \eqref{eq:pi_p_v}.
%%

@@

In analogy to Eq. \eqref{eq:ET_average},
%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:xt2_sum},
we can write the error after a period, $T$,
as a sum of two contributions,
%
\begin{equation}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T),
  \label{eq:error_tot}
\end{equation}
%
where the residual error
(from the decay of the initial error)
is given by
%
\begin{equation}
  \Err_R(T)
  =
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) },
  \label{eq:error_res}
\end{equation}
%
and the asymptotic error
is given by
%
\begin{align}
  \Err_A(T)
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot y_k^2\bigl( q(t) \bigr) \, dt
  ,
\label{eq:error_asym}
\end{align}
%
where
%
\begin{equation}
  y_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  ,
  \label{eq:uk_def}
\end{equation}
%
and $\Gamma_k$ is computed from
the autocorrelation function
of the histogram fluctuation,
$\kappa_k(t) \equiv \bigl\langle {\tilde \zeta}_k(0)
\, {\tilde \zeta}_k(t) \bigr\rangle$,
as:
%
\begin{equation}
  \Gamma_k
  = \sum_{t = -\infty}^\infty \kappa_k(t)
  = \sum_{t = -\infty}^\infty
  \left\langle {\tilde \zeta}_k(0)
            \, {\tilde \zeta}_k(t) \right\rangle
  .
  \label{eq:Gamma_sum}
\end{equation}
%
%\begin{equation}
%$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
%  \notag
%\label{eq:Gamma_integral}
%\end{equation}
%
Equation \eqref{eq:Gamma_sum} provides a way
of approximately modeling
a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
In Appendix \ref{sec:Gamma_measure},
we discuss a way of estimating $\Gamma_k$
from an FDS simulation under constant magnitude.
%

@@

\item \label{step:prerun}
Run a preliminary adaptive FDS simulation
under constant updating magnitude, $a_0$.
%
The updating magnitude should be sufficiently small
such that the resulting error is no greater than
multiple of $1.0$.

@@

We used the Metropolis algorithm\cite{frenkel, metropolis1953} for configuration sampling.
%
Each step consists of
one MC trial for displacing one of the two special particles,
and another four trials for displacing the other particles.
%
In each trial, the displacement along each dimension is uniform,
with the size given by $0.1/\varrho^*$ to ensure the acceptance ratio is around 50\%,
where $\varrho^*$ is the reduced number density.

The PMF was calculated from $r = 0$ to half of the box dimension
on a grid of size $\Delta r = 0.01$ in reduced units.
%
Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
and the numbers of bins were $n = 500$ and $250$, respectively.
%
We tested the single-bin, Gaussian, and bandpass updating schemes.
%
Each simulation started with an ``equilibration'' phase
of $10^7$ steps under constant updating magnitude,
$a_0 = 10^{-4}$,\footnote{The equivalent $\ln f = n \, a_0$ for the WL algorithm
was $0.05$ and $0.025$, respectively.}
followed by a production phase
with updating magnitude given by
either the inverse-time or optimal schedule.
%
We repeated each simulation multiple times
and report the average.
%
To make sure the same schedule for multiple runs under the same condition,
the optimal schedule was pre-computed from a set of fixed values of
$\Gamma_k$ and $\epsilon_k$,
although they can also be fairly accurately estimated from the data
collected from the equilibration phase of each individual run.

@@

In Fig. \ref{fig:lj_xerr},
we show the error components against the eigenmode index
for the Gaussian updating scheme of width $\sigma = 0.2$.
%
While the Gaussian updating scheme effectively suppressed
the initial errors for short-range or large-$k$ modes,
it also slowed down the reduction of errors of these modes.
%
As a result, the final errors of the long-range or small-$k$ modes
were larger than those from the single-bin scheme.
%
For the Gaussian updating scheme,
the inverse-time schedule tended to reduce the updating magnitude
too quickly
leaving a hump on the error profile for the middle-range eigenmodes.
%
The optimal schedule improved over the inverse-time schedule
and produced a flatter profile for the final error.
%
It was also possible to roughly
recover the single-bin result by using the correction formula
based on the histogram, Eq. \eqref{eq:UHcorr}.
%
This correction usually benefited the small-$k$ modes
but not necessarily for the large-$k$ modes.

In Fig. \ref{fig:lj_alpha}. 
we show the optimal schedules in a few cases.
%
Generally, the optimal schedule
for the $\varrho^* = 0.8$ case
was closer to the inverse-time schedule
than that for the $\varrho^* = 0.1$ case.
%
This was probably because in the former case
the first mode dominates the total error,
$\Gamma_1 \approx \Gamma$,
and the schedule was roughly optimized for this mode,
resulting in
$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
as $\lambda_1 \approx 1$
[cf. Eq. \eqref{eq:lambda_Gaussian}].
%
Since we have chosen a long equilibration period
for the preliminary run under constant updating magnitude,
the starting magnitude was usually around $a_0/2 = 5\times10^{-5}$
[cf. Eq. \eqref{eq:half_alpha0}].
%
But if the Gaussian were too wide,
e.g. if $\sigma = 0.6$,
the eigenvalues $\lambda_k$ of some modes
became too small to damp out the initial bias,
resulting in much larger initial updating magnitude, $\alpha(0)$
[cf. Eq. \eqref{eq:opt_alpha0}].
%
In this case, our method would fail
because the resulting bias potential would no longer be small,

\begin{table}[h]\footnotesize
  \caption{\label{tab:error_LJ}
    Error of the LJ system.
    The numbers have been averaged over multiple runs,
    and the predicted values are shown in parentheses for comparison;
    $\Err_\mathrm{init.}$
    denotes the error
    at the end of the equilibration phase under constant magnitude;
    $\Err_\mathrm{final}$
    denotes the error at the end of the production phase under
    either the inverse-time ($1/t$) or optimal schedule (opt.).
    \note{The reference values can be found
    on the second line of \texttt{verr.log}.}%
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Sched. &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    \hline
    \multicolumn{4}{c}{
      $\varrho^* = 0.1$,
      $500$ bins,
      $\Gamma_1 \approx 111$,
      $\Gamma \approx 1.1\times10^3$.
    } \\
    \hline
    single-bin & $1/t$
    & $5.7(5.7)\times10^{-2}$
    & $9.1(11.4)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $7.7(8.9)\times10^{-3}$
    & $6.8(8.2)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $7.6(8.9)\times10^{-3}$
    & $3.2(3.9)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $6.5(7.8)\times10^{-3}$
    & $4.0(4.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $7.0(7.8)\times10^{-3}$
    & $2.5(2.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $5.5(6.6)\times10^{-3}$
    & $3.2(3.8)\times10^{-5}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $5.8(6.6)\times10^{-3}$
    & $2.6(3.0)\times10^{-5}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $8.0(8.9)\times10^{-3}$
    & $1.6(1.8)\times10^{-5}$
    \\
    \hline
    \multicolumn{4}{c}{
      $\varrho^* = 0.8$,
      $250$ bins,
      $\Gamma_1 \approx 2.0\times10^4$,
      $\Gamma \approx 2.7\times10^4$
    } \\
    \hline
    single-bin & $1/t$
    & $1.3(1.4)$
    & $2.7(2.7)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & $1/t$
    & $1.3(1.3)$
    & $2.9(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.1$ & opt.
    & $1.2(1.3)$
    & $2.8(2.8)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & $1/t$
    & $1.3(1.2)$
    & $3.4(3.2)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.2$ & opt.
    & $1.2(1.2)$
    & $3.1(2.9)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & $1/t$
    & $0.9(0.9)$
    & $9.8(8.0)\times10^{-3}$
    \\
    %\hline
    $\sigma=0.5$ & opt.
    & $1.0(0.9)$
    & $5.8(5.4)\times10^{-3}$
    \\
    %\hline
    $K=20$ & $1/t$
    & $1.3(1.3)$
    & $3.1(2.7)\times10^{-3}$
    \\
    \hline
  \end{tabular}
\end{table}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_xerr.pdf}
  }
  \caption{
    \label{fig:lj_xerr}
    Error components for the FDS simulation on the LJ system
    using the Gaussian updating scheme of width $\sigma = 0.2$.
    %
    The optimal schedule (opt.) is more effective
    in reducing the errors of middle-range eigenmodes
    than the inverse-time schedule ($1/t$).
    The lines are a guide to the eyes.
  }
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_alpha.pdf}
  }
  \caption{
    \label{fig:lj_alpha}
    Optimal schedule of the Gaussian updating scheme
    for the LJ system.
    %
    The starting updating magnitude was usually
    around $a_0/2 = 5\times10^{-5}$
    as given by Eq. \eqref{eq:half_alpha0}.
    But an overly wide Gaussian window ($\sigma = 0.6$)
    led to a much larger updating magnitude.
  }
\end{center}
\end{figure}

@@

To begin, we split the entire simulation into $S$ stages,
each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
%
We denote the number of steps in stage $s$ as $T_s$, and
$\sum_{s=1}^S T_s = T$.
%
%\begin{equation}
%  \sum_{s = 1}^S T_s = T
%  ,
%  \label{eq:sumT}
%\end{equation}
%
We assume that
the magnitude follows geometrically,
%
\begin{equation}
  \alpha_s = \alpha_0 \, r^s
  ,
  \label{eq:alphas}
\end{equation}
%
with $0 < r < 1$.
%
%and that the system is properly equilibrated
%under constant magnitude initially.
%
We wish to find the best distribution of the simulation time
into the $S$ stages
as well as
the optimal number of stages and $r$.

@@

Here we give a method of measuring
the integrals of the autocorrelation functions, $\Gamma_k$'s,
in an adaptive FDS simulation
under constant updating magnitude.
%
This method has the advantage of not requiring
explicit computation of 
the autocorrelation functions of the eigenmodes.
%
The basic idea is to use Eq. \eqref{eq:xt2_eql}.
%
However,
without knowing the intrinsic distribution, $p^*_i$,
we cannot compute the shifted bias potential, $v_i$.
%
The problem can be avoided by replacing the shift
by the time average of the bias potential.

We first note that the difference, 
$u_i(t) - v_i(t) = \ln(p^*_i/p_i)$, is a constant of time,
so is the Fourier transform,
$\tilde u_k(t) - \tilde v_k(t)$.
%
It follows that
${\tilde u}_{k}$ and ${\tilde v}_{k}$ would also share the same variance.
%
But since the ensemble average of ${\tilde v}_{k}$
is zero, we have
\begin{equation*}
  \operatorname{var} {\tilde u}_{k}
  =
  \operatorname{var} {\tilde v}_{k}
  =
  \left\langle {\tilde v}_{k}^2 \right\rangle
  .
\end{equation*}
%
Finally, by using Eq. \eqref{eq:xt2_eql},
we get
%
\begin{equation}
  \operatorname{var} {\tilde u}_k
  =
  \frac{1}{2} \,
  a_0 \, \Gamma_k \, \lambda_k,
\label{eq:varXt}
\end{equation}
%
where the variance can be computed from trajectory averages.
%
A practical limitation is that Eq. \eqref{eq:varXt}
would fail if some eigenvalue, $\lambda_k$, is close to zero.
So it is best used for the single-bin updating scheme.

@@

For the Gaussian updating kernel,
if the width of the angular variable, $\sigma_\varphi$,
satisfies $\Delta \varphi \ll \sigma_\varphi \ll \pi$,
we can extend
the upper limit of the integrals
to infinity, and
%
\begin{equation}
  \mu(\varphi)
  =
  \frac{            1            }
       { \sqrt{ 2 \pi \sigma_\varphi^2 } }
  %
  \exp\left(
        -\frac{   \varphi^2   }
              { 2 \, \sigma_\varphi^2 }
      \right)
  .
\notag
\end{equation}
%
%
It follows that all eigenvalues are positive\cite{bussi2006}:
%
\begin{align}
  \lambda_k
  &=
  \exp\left(
        -\frac{ k^2 \, \sigma_\varphi^2 }
              {           2             }
      \right)
  =
  \exp\left[
        -\frac 1 2
        \left(
         \frac{ 2 \, \pi \, k \, \sigma_z }
              { g \, n \, \Delta z  }
        \right)^2
      \right]
  ,
%\notag
\label{eq:lambda_Gaussian}
\end{align}
where $\sigma_z$ and $\Delta z$
are the width of the Gaussian
and the bin width
in the original collective variable, $z$,
respectively.
